---
layout: post
title: "ALM in R"
category: examples
tags: [r]
authors:
 - surname: Chamberlain
   given-names: Scott
   orcid: http://orcid.org/0000-0003-1444-9135
   email: scott@ropensci.org
---

The ALM application is now being used by other publishers. Since other publishers use this same application, we can access altmetrics data from all of them with a single client. rOpenSci has made an R client for interacting with the ALM application from R, called `alm`. 

There are now four publishers that are using the ALM app - that is, that you can get altmetrics data from: PLOS, Crossref, Publick Knowledge Project (PKP), and Copernicus. More are likely coming on line.

As a quick demonstration of the power of this single interface to many different altmetrics providers, below are a few examples. 

### Install and load alm

```{r eval=FALSE}
install_github("ropensci/alm")
```

```{r messages=FALSE}
library("alm")
```


### Get data

#### PLOS article data

The default in the `alm` package is for the PLOS ALM app. You do need to get an API key first here [http://alm.plos.org/](http://alm.plos.org/). You can pass in the `key` parameter or store in your `.Rprofile` file and pass in that way, or do `options(PlosApiKey = "yourkey")` and that will be stored for your current R session.


```{r}
out <- alm(doi = "10.1371/journal.pone.0036240")
head(out)
```

#### Crossref article data

You need to get a Crossref ALM API key first here [http://alm.labs.crossref.org/docs/Home](http://alm.labs.crossref.org/docs/Home), and pass in a different URL


```{r}
url <- "http://alm.labs.crossref.org/api/v3/articles"
alm(doi = "10.1080/15459624.2013.816432", url = url, key = getOption("crossrefalmkey"))
```


#### Public Knowledge Project (PKP) article data

You need to get a PKP ALM API key first here [http://pkp-alm.lib.sfu.ca/](http://pkp-alm.lib.sfu.ca/), and pass in a different URL


```{r}
url <- "http://pkp-alm.lib.sfu.ca/api/v3/articles"
alm(doi = "10.3402/gha.v7.23554", url = url, key = getOption("pkpalmkey"))
```


#### Copernicus publishers article data

You need to get a Copernicus ALM API key first here [http://metricus.copernicus.org/](http://metricus.copernicus.org/), and pass in a different URL


```{r}
url <- "http://metricus.copernicus.org/api/v3/articles"
alm(doi = "10.5194/fr-17-33-2014", url = url, key = getOption("copernicusalmkey"))
```

<hr>

Once you have data you can leverage the great statistics and visualization power of R easily.